Variational Monte Carlo on Bosonic Systems

### 2018-04-06 project_1/article/article.tex
Spell check gradient descent section

        \subsection{Gradient Descent}

            Scanning the entire space of variational parameters can be a
            somewhat despairing task. To aid us in the search we therefore
            implement the method of gradient descent\footnote{Also known as
            steepest descent, easily confused with the steepest descent method
            for integation.}. In any optimum we generally have that the
            derivative should equal zero. In our case, we want to find the
            minimum of the expected local energy with respect to some
            variational parameter $\alpha$, so that
            \begin{equation}
                \nabla_\alpha \expval{E_L(\mathbf{R}; \alpha)} = 0.
            \end{equation}
            The foundation for  the method of gradient descent lies in the fact
            that some analytic \footnote{We use the term analytic quite
            recklessly.} function $\vf{F}(\alpha)$ decreases fastest in the direction
            of $-\nabla_\alpha \cdot \vf{F}(\alpha)$. This means that we could eventually
            get to an ensuing minimum by trailing along the pathway of the
            following recursive relation,
            \begin{equation}
                \vfg{\alpha}_{k+1}
                = \vfg{\alpha}_k - \gamma \nabla_{\alpha}\cdot \vf{F}(\alpha_k).
                \label{eq:parameter_difference}
            \end{equation}
            If $\gamma$ is small enough we should have $\vf{F}(\alpha_k) \geq
            \vf{F}(\alpha_{k+1})$ for all $k$ and hopefully we will experience
            convergence to the desired minimum.

            For our special case we only have one variational parameter,
            $\alpha$.  The expression for the derivative is,
            \begin{align}
                \nabla_{\alpha}\expval{E_L}
                =
                2\left(
                    \expval{E_L\frac{1}{\Psi_T}\nabla_{\alpha}\Psi_T}
                    - \expval{\frac{1}{\Psi_T} \nabla_{\alpha}\Psi_T}
                    \expval{E_L}
                \right).
                \label{eq:variational_parameter_gradient}
            \end{align}

            To evaluate this expression we sample the local energy, the
            parameter gradient of the wavefunction, and the combined expression
             of these at each Monte Carlo cycle.

@@

\section{Variational parameter gradient of the expectation energy}
    Here we show how to arrive at the expression shown in
    \autoref{eq:variational_parameter_gradient}. This requires us to restrict
    our view to real trial wavefunctions, i.e., $\Psi_T^{*} = \Psi_T$, which is
    the case for the wavefunctions we are exploring.
    \begin{align}
        \nabla_{\alpha}\expval{E_L}
        &=
        \nabla_{\alpha}
        \para{
            \frac{
                \int\dd\vf{r}\Psi_T^{*}H\Psi_T
            }{
                \int\dd\vf{r}|\Psi_T|^2
            }
        }
        \\
        &=
        -
        \frac{
            2
            \para{
                \int\dd\vf{r}\Psi_T H\Psi_T
            }
        }{\para{\int\dd\vf{r}\Psi_T}^2}
        \int\dd\vf{r}\Psi_T\nabla_{\alpha}\Psi_T
        \\
        &\qquad
        +
        \frac{2}{\int\dd\vf{r}\Psi_T^2}
        \int\dd\vf{r}\Psi_TH\brak{
            \nabla_{\alpha}\Psi_T
        }
        \\
        &=
        - 2\expval{E_L}
        \frac{1}{
            \int\dd\vf{r}\Psi_T^2
        }
        \int\dd\vf{r}\Psi_T^2\para{
            \frac{\nabla_{\alpha}\Psi_T}{\Psi_T}
        }
        \\
        &\qquad
        +
        \frac{2}{\int\dd\vf{r}\Psi_T^2}
        \int\dd\vf{r}\Psi_T^2 E_L\para{
            \frac{\nabla_{\alpha}\Psi_T}{\Psi_T}
        }.
    \end{align}
    We now use the definition of the expectation value to group terms together.
    We are thus left with
    \begin{align}
        \nabla_{\alpha}\expval{E_L}
        =
        2
        \para{
            \expval{
                E_L\frac{1}{\Psi_T}\nabla_{\alpha}\Psi_T
            }
            - \expval{E_L}
            \expval{\frac{1}{\Psi_T}\nabla_{\alpha}\Psi_T}
        },
    \end{align}
    which is what we wanted to show.

>>>>

        \subsection{Gradient Descent}
            Scanning the entire space of variational parameters can be a
            somewhat despairing task. To aid us in the search we therefore
            implement the method of gradient descent\footnote{Also known as
            steepest descent, easily confused with the steepest descent method
            for integation.}. In any optimum we generally have that the
            derivative should equal zero. Our goal is to find the minimum of the
            expected local energy with respect to the variational parameters
            $\vfg{\alpha}$.
            \begin{equation}
                \nabla_{\vfg{\alpha}} \expval{E_L(\mathbf{R}; \vfg{\alpha})}
                = 0.
            \end{equation}
            The foundation for  the method of gradient descent lies in the fact
            that some analytic \footnote{We use the term analytic quite
            recklessly.} function $\vf{F}(\vfg{\alpha})$ decreases fastest in
            the direction of $-\nabla_{\vfg{\alpha}} \cdot
            \vf{F}(\vfg{\alpha})$. This means that we could eventually get to an
            ensuing minimum by trailing along the pathway of the following
            recursive relation,
            \begin{equation}
                \vfg{\alpha}_{k+1}
                = \vfg{\alpha}_k
                - \gamma \nabla_{\vfg{\alpha}}\cdot \vf{F}(\vfg{\alpha}_k).
                \label{eq:parameter_difference}
            \end{equation}
            If $\gamma$ is small enough we should have $\vf{F}(\vfg{\alpha}_k)
            \geq \vf{F}(\vfg{\alpha}_{k+1})$ for all $k$ and hopefully we will
            experience convergence to the desired minimum.

            The variational gradient of the expectation value of the energy is
            thus
            \begin{align}
                \nabla_{\vfg{\alpha}}\expval{E_L}
                =
                2\left(
                    \expval{E_L\frac{1}{\Psi_T}\nabla_{\vfg{\alpha}}\Psi_T}
                    - \expval{\frac{1}{\Psi_T} \nabla_{\vfg{\alpha}}\Psi_T}
                    \expval{E_L}
                \right).
                \label{eq:variational_parameter_gradient}
            \end{align}
            To evaluate this expression we sample the local energy, the
            parameter gradient of the wavefunction, and the combined expression
            of these at each Monte Carlo cycle.

@@

\section{Variational parameter gradient of the expectation energy}
    Here we show how to arrive at the expression shown in
    \autoref{eq:variational_parameter_gradient}. This requires us to restrict
    our view to real trial wavefunctions, i.e., $\Psi_T^{*} = \Psi_T$, which is
    the case for the wavefunctions we are exploring.
    \begin{align}
        \nabla_{\vfg{\alpha}}\expval{E_L}
        &=
        \nabla_{\vfg{\alpha}}
        \para{
            \frac{
                \int\dd\vf{r}\Psi_T^{*}H\Psi_T
            }{
                \int\dd\vf{r}|\Psi_T|^2
            }
        }
        \\
        &=
        -
        \frac{
            2
            \para{
                \int\dd\vf{r}\Psi_T H\Psi_T
            }
        }{\para{\int\dd\vf{r}\Psi_T}^2}
        \int\dd\vf{r}\Psi_T\nabla_{\vfg{\alpha}}\Psi_T
        \\
        &\qquad
        +
        \frac{2}{\int\dd\vf{r}\Psi_T^2}
        \int\dd\vf{r}\Psi_TH\brak{
            \nabla_{\vfg{\alpha}}\Psi_T
        }
        \\
        &=
        - 2\expval{E_L}
        \frac{1}{
            \int\dd\vf{r}\Psi_T^2
        }
        \int\dd\vf{r}\Psi_T^2\para{
            \frac{\nabla_{\vfg{\alpha}}\Psi_T}{\Psi_T}
        }
        \\
        &\qquad
        +
        \frac{2}{\int\dd\vf{r}\Psi_T^2}
        \int\dd\vf{r}\Psi_T^2 E_L\para{
            \frac{\nabla_{\vfg{\alpha}}\Psi_T}{\Psi_T}
        }.
    \end{align}
    We now use the definition of the expectation value to group terms together.
    We are thus left with
    \begin{align}
        \nabla_{\vfg{\alpha}}\expval{E_L}
        =
        2
        \para{
            \expval{
                E_L\frac{1}{\Psi_T}\nabla_{\vfg{\alpha}}\Psi_T
            }
            - \expval{E_L}
            \expval{\frac{1}{\Psi_T}\nabla_{\vfg{\alpha}}\Psi_T}
        },
    \end{align}
    which is what we wanted to show.
